---
title: "p8105_hw1_xg2451"
author: "Xuanyu Guo"
date: "2024-09-17"
output: html_document
---
# Problem 1
```{r}
# load the penguins dataset
data("penguins", package = "palmerpenguins")
```
The `penguins` dataset includes several variables about penguin measurements and observations such as species, island, bill length (mm), bill depth (mm), flipper length (mm), body mass (g), sex and study year.

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length across all penguins in the dataset is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

```{r}
# Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable
library(ggplot2)
penguins <- na.omit(penguins)  # Remove rows containing missing values

p <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Flipper Length vs Bill Length",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)")
p
```
```{r}
# Export scatterplot to project directory using ggsave
ggsave("penguins_scatterplot.png", plot = p, width = 10, height = 8)
```















